---
title: "Notional Data Exploration"
format: 
  html:
    toc: true
    df-print: kable
execute-dir: project
execute:
  echo: false
  warning: false
---

```{r}
#| include: false

library(sf)
library(tidyverse)
library(mapview)
library(kableExtra)
library(glue)

theme_set(theme_minimal())
setwd(here::here())

input <- read_csv('data/v1/clad_omop_degauss_simulated_input_old.csv') %>%
  rename(Location_id = id) %>%
  mutate(id = row_number() - 1) %>%
  inner_join(
    read_csv('data/v1/OMOP_sample_simulated_residential.csv'), 
    by = 'Location_id'
  )
raw_output <- list(
  postgis = readxl::read_excel('data/v1/clad_omop_simulated_output_4.xlsx', sheet = 'postgis_responses', col_types = 'text'),
  degauss = readxl::read_excel('data/v1/clad_omop_simulated_output_4.xlsx', sheet = 'degauss_responses', col_types = 'text'),
  nominatim = readxl::read_excel('data/v1/clad_omop_simulated_output_4.xlsx', sheet = 'nominatim_responses', col_types = 'text')
) %>% map(select, -c(matches('json')))

output <- list(
  postgis = readxl::read_excel('data/v1/clad_omop_simulated_output_4.xlsx', sheet = 'postgis_responses') %>% 
    transmute(
      id, 
      rating_postgis = rating,
      lat_gc = geo_lat,
      long_gc = geo_long
    ) %>%
    group_by(id) %>%
    mutate(result_postgis = row_number()),
    degauss = readxl::read_excel('data/v1/clad_omop_simulated_output_4.xlsx', sheet = 'degauss_responses') %>% 
    transmute(
      id, 
      score_degauss = geo_score,
      precision_degauss = geo_precision,
      lat_gc = geo_lat,
      long_gc = lon
    ) %>%
    group_by(id) %>%
    mutate(result_degauss = row_number()) %>%
    filter(precision_degauss != 'zip'),
    nominatim = readxl::read_excel('data/v1/clad_omop_simulated_output_4.xlsx', sheet = 'nominatim_responses')  %>% 
    transmute(
      id, 
      type_gc = geo_type,
      lat_gc = as.numeric(geo_lat),
      long_gc = as.numeric(geo_long),
    ) %>%
    group_by(id) %>%
    mutate(result_nominatim = row_number())
) %>%
  map(st_as_sf, coords = c('long_gc', 'lat_gc'), crs = 4326, remove = FALSE, na.fail = FALSE)

output_long <- output %>% 
  bind_rows(.id = "geocoder") 
compare <- input %>%
  select(
    id, 
    state_abbr, 
    address_type, 
    location_source_value, 
    lat_input = latitude, 
    long_input = longitude
  ) %>% 
  expand_grid(geocoder = c('postgis', 'degauss', 'nominatim')) %>% 
  left_join(output_long, by = c('id', 'geocoder')) %>% 
  mutate(failure = is.na(lat_gc) | is.na(long_gc)) %>% 
  arrange(id, geocoder)

compare$geometry_true <- st_as_sf(as.data.frame(compare), coords = c("long_input","lat_input"), crs = 4326)$geometry

compare <- compare %>% mutate(
    gc_diff_m = as.numeric(st_distance(geometry, geometry_true, by_element = TRUE)),
    location_source_value = toupper(location_source_value)
  ) 

write_csv(compare, 'data/v1/geocode_comparison_clean.csv')
```

## Data Description

### Input

The input data set of 523 addresses contains a variety of public places with verified locations.

The output is indexed by the row order of the file that was input into it. That is the `id` field shown in any of these outputs.

##### Sample Input

```{r}
head(st_drop_geometry(input))
```

### Output

Input data was geocoded with three different geocoders.

#### PostGIS

##### Notes

-   Based on US Census TIGER database

-   Contains a "rating" column indicating match confidence.

-   Documentation [here](https://postgis.net/docs/manual-3.4/postgis_installation.html#loading_extras_tiger_geocoder)

##### Sample Output

```{r}
head(raw_output$postgis)
```

#### Degauss

##### Notes

-   Contains a "score" rating the confidence in the match, and a "precision" stating which kind of match was found ("street", "range", or "zip"). All "zip" matches were censored.

-   Documentation [here](https://degauss.org/geocoder/)

##### Sample Output

```{r}

head(raw_output$degauss)
```

#### Nominatim

##### Notes

-   Based on OpenStreetMap database

-   Documentation [here](https://nominatim.org/release-docs/latest/)

##### Sample Output

```{r}
head(raw_output$nominatim)
```

## Performance by Geocoder

Each geocoder was given 523 input addresses. 

Degauss and PostGIS performed notably better than Nominatim on both accuracy and success rate. Nominatim failed to return a match almost a quarter of the time, and when it did, it often returned several matches.

```{r}
gc_summary <- compare %>%
  group_by(geocoder, id) %>%
  summarize(
    gc_diff_m = mean(gc_diff_m),
    address_gc_count = n(),
    all_fail = all(isTRUE(failure))
  ) %>%
  ungroup() %>% 
  group_by(geocoder) %>%
  summarize(
    `Min` = min(gc_diff_m, na.rm=TRUE), 
    `Mean` = mean(gc_diff_m, na.rm=TRUE), 
    `Median` = median(gc_diff_m, na.rm=TRUE), 
    `Max` = max(gc_diff_m, na.rm=TRUE), 
    `Multiple` =  glue('{sum(address_gc_count > 1)} ({round(100*sum(address_gc_count > 1)/n())}%)'),
    `None` = glue('{sum(all_fail)} ({round(100*sum(all_fail)/n())}%)') 
  ) %>%
  arrange(geocoder) 
gc_summary %>%
  kbl(
    digits=0, 
  ) |> 
  kable_material() %>% 
  add_header_above(
    c(" ",
      "Distance from Gold Standard (m)" = 4,
      "Matches" = 2
    ),
    align = "c"
    )
  
```

```{r}
ggplot(compare) + 
  geom_boxplot(aes(x = geocoder, y = gc_diff_m)) +
  ylim(0, 1000) + 
  ylab('Distance from Gold Standard (m) - Truncated') +   
  xlab('Geocoder') +
  ggtitle("Error by Geocoder", subtitle = "Error Truncated to 1000m")
```

## Geocode Confidence Ratings

Two of the geocoders, PostGIS, and Degauss, provide some diagnostic information about their confidence in the geocode. We would like to know if those are actually predictive of the geocode's accuracy.

### PostGIS

PostGIS provides a numerical rating from 0 to 100, with zero being the most confident.

Because the distribution of geocoding errors has a long tail, we'll look at the relationship between the rating and the error on both the natural and logarithmic scale. The natural scale is truncated to an error of 1000m in order to see the differences happening at that scale.

```{r}
ggplot(filter(compare, geocoder == 'postgis')) + 
  geom_point(aes(x = rating_postgis, y = gc_diff_m)) +
  ylim(0, 1000) + 
  ylab('Distance from Gold Standard (m) - Truncated') +   
  xlab('Geocode Rating') +
  ggtitle("Error vs. Geocode Rating", subtitle = "Error Truncated to 1000 m")
```

```{r}
ggplot(filter(compare, geocoder == 'postgis')) + 
  geom_point(aes(x = rating_postgis, y = gc_diff_m)) +
  scale_y_continuous(trans = 'log10') +
  ylab('Distance from Gold Standard (m)') +   
  xlab('Geocode Rating') + 
    ggtitle("Error vs. Geocode Rating - Log Scale")
```

There is not an obvious visual relationship between the geocoding error and the geocode rating. A cursory web search did not reveal a definition for the rating.

### Degauss

Degauss provides a "precision" value ("street", "range", or "zip" in the notional data) and a "score" value. Scores are only comparable within precision groups per the Degauss docs. They define the score as:

> The percentage of text match between the given address and the geocoded result, expressed as a number between 0 and 1. A higher score indicates a closer match.

Again, we'll look at the relationship between the rating and the error on both the natural and logarithmic scale, this time broken out by the geocode precision.

```{r}
ggplot(filter(compare, geocoder == 'degauss' & precision_degauss != 'zip')) + 
  geom_point(aes(x = score_degauss, y = gc_diff_m)) +
  ylim(0, 1000) + 
  facet_wrap(~precision_degauss) +   
  ylab('Distance from Gold Standard (m) - Truncated') +   
  xlab('Geocode Score') + 
  ggtitle("Error vs. Geocode Score", subtitle = "Truncated to 1000 m")
```

```{r}
ggplot(filter(compare, geocoder == 'degauss' & precision_degauss != 'zip')) + 
  geom_point(aes(x = score_degauss, y = gc_diff_m)) +
  scale_y_continuous(trans = 'log10') + 
  facet_wrap(~precision_degauss) + 
  ylab('Distance from Gold Standard (m)') +   
  xlab('Geocode Score') + 
  ggtitle("Error vs. Geocode Score - Log Scale")
```

Degauss's geocode score does also not seem to be extremely predictive of the geocoding error. However, "street" precision does have more error than "range" precision, as would be expected. There does appear to be a relationship on the log scale between geocode score and geocode error within the "street" precision type (r² = 0.29), but not a very strong one. 


### Nominatim

Nominatim results did not include any sort of confidence score to evaluate.

## Multiple Results

PostGIS only ever returned a single response.

Degauss returns multiple results when their geocode score is tied. This occurred in 9 locations.

Nominatim is much more likely to return multiple results, and does not provide an obvious way to rank them.

### Degauss

Of the nine addresses for which Degauss returned multiple tied geocodes, only one address had more than two geocodes. Since they are tied, there's not an obvious way to pick between them.

Setting aside the single case where Degauss returned more than two results, let's look at how the geocodes relate to each other. I also include the geocoding error for each location, the geocoding error for the centroid of the two locations, and the geocoding error of PostGIS for comparison. 

```{r}
opts <- options(knitr.kable.NA = "No Match")
compare %>%
  filter(geocoder == 'degauss') %>%
  group_by(geocoder, id) %>%
  filter(n() == 2) %>%
  transmute(id, location_source_value, geometry, geometry_true, i = paste0('geometry_', row_number())) %>%
  pivot_wider(names_from = i, values_from = geometry) %>%
  ungroup() %>% 
  left_join(filter(compare, geocoder == 'postgis') %>% select(id, `PostGIS Geocode Error (m)` = gc_diff_m), by = 'id') %>% 
  transmute(
    id,
    `Address` = location_source_value,
    `Distance between Geocodes (m)` = as.numeric(st_distance(geometry_1, geometry_2, by_element = TRUE)),
    `Geocode 1 Error (m)` = as.numeric(st_distance(geometry_1, geometry_true, by_element = TRUE)), 
    `Geocode 2 Error` = as.numeric(st_distance(geometry_2, geometry_true, by_element = TRUE)), 
    `Centroid Error (m)` = as.numeric(st_distance(
      st_centroid(
        st_union(geometry_1, geometry_2)
      ), 
      geometry_true, 
      by_element = TRUE)),
    `PostGIS Geocode Error (m)`
  ) %>% 
  kbl(
    digits=0
  ) 
    
```

It seems reasonable to either pick the centroid of ties, or pick one at random. Or a tied result could be a cue for us to fall back on another geocoder.

### Nominatim

As mentioned above, Nominatim returned multiple results a whopping 25% of the time, and it does not provide any way to rank their accuracy.

Here are the three worst cases of Nominatim geocodes (as measured by the area of by the polygon they form). The "true" location is shown as a large blue circle.

```{r}
options(opts)

baddies <- output_long %>% 
  filter(geocoder == 'nominatim') %>% 
  group_by(geocoder, id) %>% 
  summarize(geometry = st_union(geometry)) %>% 
  ungroup() %>% 
  st_as_sf() %>% 
  st_convex_hull() %>% 
  filter(as.numeric(st_area(geometry)) > 0) %>% 
  arrange(desc(st_area(geometry))) %>% 
  head(n=3) %>% 
  inner_join(input, by = 'id')

baddies <- compare %>% 
  filter(id %in% baddies$id) %>%
  group_by(id) %>%
  group_split()
```

#### `r unique(baddies[[1]]$location_source_value)`

In this case, Nominatim may have interpretted "Providence" as the county name. All these locations are on streets named "Dexter" in Providence County. Both PostGIS and Degauss get quite close.

```{r}
mapview(baddies[[1]]$geometry_true, cex = 10) + 
mapview(st_as_sf(baddies[[1]]), zcol = 'geocoder', layer.name = 'Geocoder') 
```

#### `r unique(baddies[[2]]$location_source_value)`

This may be a epistemologically-challenging case of the "truth" being incorrect. However, assuming that Degauss and PostGIS are getting close to the "true truth", we still see Nominatim returning several guesses along a length of road. 

```{r}
mapview(baddies[[2]]$geometry_true, cex = 10) + 
mapview(st_as_sf(baddies[[2]]), zcol = 'geocoder', layer.name = 'Geocoder') 
```

#### `r unique(baddies[[3]]$location_source_value)`

Here, Nominatim made several guesses along the correct roadway. Both PostGIS and Degauss got reasonably close. 

```{r}
mapview(baddies[[3]]$geometry_true, cex = 10) + 
mapview(st_as_sf(baddies[[3]]), zcol = 'geocoder', layer.name = 'Geocoder') 
```
